 ---
title: "Assignment 1 Submission"
author: "Ian Chen, Marco Gunawan"
date: "submission date"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Statement of Contribution

Include this if pairing.

State what each team member has contributed to the assignment before answering the questions. Each student must contribute towards ALL questions. Your assignment will not be marked without this statement.

```{r eval=FALSE, include=FALSE}
## There are a few options in every chunk, click the setting button (greyed) near the triangle:
# 1. to use default as in the above (show code and output, ie ECHO=TRUE) or 
# 2. to show output only (ie ECHO=FALSE) or
# 3. show nothing (run code) ie include=FALSE
# 4. show nothing (dont run the code), like this chunk, ie eval=FALSE, include=FALSE
## You can knit this template into Word and then update it into pdf etc..
## You can create your own way for submission, this is just an example if you are going to use Rmd.
```

# Question 1 Normal distribution. (8 marks)

(a) (1 mark) Find the probability that the user spends more than 15 minutes per month at the site.

    **Answer: 0.9937903**

    *Answer text here: *

```{r echo=TRUE}
    pnorm(15,mean = 25, sd = 4,lower.tail=FALSE)
    ```

(b) (2 marks) Find the probability that the user spends between 20 and 35 minutes per month at the site.

    **Answer:0.8881406**

    *words words blah blah blah thingy minus thingy*

    ```{r include=FALSE}
    pnorm(35,25,4) - pnorm(20,25,4)
    ```

(c) (2 marks) What is the amount of time per month a user spends on Facebook, if only 1% of users spend this time or longer on Facebook?

    **Answer: 34.30539 **

    *Answer text here*
    
```{r}
qnorm(0.99,25,4)
```
    

(d) (3 marks) Between what values do the time spent of the middle 90% distribution of Facebook users fall?

    **Answer: 18.42059 and 31.57941**

    *Answer text here*
    
```{r}
qnorm(c(0.05, 0.95), 25, 4)
```
    
```{r}
qnorm(0.05, 25, 4)  # 5th percentile
qnorm(0.95, 25, 4)  # 95th percentile
```

# Question 2 Blood fat concentration (11 marks)

(a) (6 marks) Conduct a two-independent sample $t$-test using R to determine whether the concentration of plasma cholesterol is significantly different between patients with no evidence of heart disease and those with narrowing of the arteries.

    **Answer: Statistically significant difference in plasma cholesterol levels between patients with and without heart disease, very small pvalue much less than 0.01**

    *Answer text here*
    
    ```{r}
# Group 1 (No disease)
mean1 <- 195.2745
var1 <- 1303.9231
sd1 <- sqrt(1303.9231)  # Convert variance to SD
n1 <- 51

# Group 2 (Disease)
mean2 <- 216.1906
var2 <- 1850.2488
sd2 <- sqrt(1850.2488)   # Convert variance to SD
n2 <- 320


t_statistic <- (mean1-mean2)/sqrt((var1/n1)+(var2/n2))

print(t_statistic)

df <- ( (var1/n1 + var2/n2)^2 ) / ( (var1/n1)^2/(n1-1) + (var2/n2)^2/(n2-1) )

print(df)

p_value <- 2 * (1 - pt(abs(t_statistic), df)) # x2 because two sided

print(p_value)


```

(b) (3 marks) Determine a 99% confidence interval for the mean difference in concentration of plasma cholesterol between the two groups of patients.

    **Answer: 99% CI: ( -35.71644 , -6.115758 )**

    *Answer text here*
    
```{r}
# Given data
mean1 <- 195.2745
var1 <- 1303.9231
n1 <- 51

mean2 <- 216.1906
var2 <- 1850.2488
n2 <- 320

# Compute standard error
SE <- sqrt((var1/n1) + (var2/n2))

df <- ( (var1/n1 + var2/n2)^2 ) / ( (var1/n1)^2/(n1-1) + (var2/n2)^2/(n2-1) )

t_critical <- qt(0.995, df)

margin_of_error = t_critical * SE

CI_lower <- (mean1 - mean2) - margin_of_error
CI_upper <- (mean1 - mean2) + margin_of_error

cat("99% CI: (", CI_lower, ",", CI_upper, ")\n")




```
    

(c) (2 marks) Explain the correspondence between the confidence interval in (b) and a test of the hypotheses you listed in question (a).

    **Answer: The Confidence interval does not include zero, hence the data suggests that we can conclude that there is a statistically difference between datasets. This corresponds with the low pvalue from a **

    *Answer text here*

# Question 3 Regression (31 marks)

(a) (2 marks) Fit a simple linear model $M_1$ to these data. Present the appropriate scatterplot and plot the fitted line onto the scatterplot. Comment about the output in a few concise sentences.

    **Answer:**
    
The output depicts a positive linear relationship between the volume of a ship's cargo (tonnage) and the time required to load and unload it. This suggests that the larger the tonnage, the longer it will take to load and unload it. However, the strength of the relationship is weak past the 7500 ton mark as there are insufficient data points in that range. Most observations fall below 7500 tons, with only two data points above this threshold. Hence this model may not be reliable in this higher tonnage range.
    
    

    *Answer text here*
    
```{r}
glakes <- read.table("glakes.txt", header = TRUE)
model_m1 <- lm(Time ~ Tonnage, data = glakes)

plot(Time ~ Tonnage, data = glakes, xlab = "Tonnage", ylab = "Time", main = "Scatterplot of Time vs Tonnage")

```
    

(b) (5 marks) Provide the model summary and diagnostics checking plots for model $M_1.$ Does the straight line regression model $M_1$ seem to fit the data well? Comment about the output in a few concise sentences.

    **Answer:**
    Question 3b redo)

Model Summary:

The fitted model equation is : Time= 12.3447+ 0.006518 × Tonnage

Slope interpretation βˆ1 = 0.006518: Time is expected to increase by 0.006518 units on average for every 1 unit increase in Tonnage.

Intercept interpretation βˆ 0 = 12.3447: This means that when the tonnage is zero, the time taken to unload and load the ship is expected to be 12.3447 units. This measurement could be the for the time the ship spends in port doing other activities such as refueling or stocking food for the crew.

The p-values for both coefficients are both extremely small, being under 0.001 which means we have strong evidence to reject the null hypothesis. Thus it is suggested that the tonnage has a significant impact on time.

Diagnostic Plots Analysis:

Residuals vs Fitted: The plot shows no clear pattern. This suggests that the relationship is approximately linear.

Q-Q Residuals: The majority of the data-points are on the diagonal line, suggesting the residuals are apporiximetly normally distrubted. However. there is some deviation towards the extremes indicating there may be some outliers within the data.

Scale-Location: The line trends slightly upwards which suggests that the variance may not be constant and is in fact increasing.

Residuals vs Leverage: The residuals vs leverage plot depicts some areas that require investigation. A large leverage means that the singular datapoint has a high influence on the output of the model. A high absolute value of the standardized residual means that they are poorly predicted by the model. Therefore the areas that require further investigation are those that have both a high leverage and a standard residual far from zero. Those values are datapoints 31 and 3.

    *Answer text here*
    
    
    
```{r}
summary(model_m1)

par(mfrow = c(2, 2))  # Arrange plots in a 2x2 grid
plot(model_m1)
par(mfrow = c(1, 1))  # Reset plot layout
```
    

(c) (5 marks) Do you think there are outliers or influential points in the data? What influence do these points have on the model fit? Use leverage and Cook's distance for this investigation.

    **Answer:**

    *Answer text here*

(d) (4 marks) Fit a regression model to the transformed $M_2$ model. Present the appropriate scatterplot and plot the fitted line onto the scatterplot. Does the transformed line regression model $M_2$ seem to fit the data well? Comment about the output in a few concise sentences.

    **Answer:**
Just like M1, the transformed line (M2) depicts a positive linear relationship between the square root of the volume of a ship's cargo (tonnage) and the logarithm of the time required to load and unload it. This suggests that as the square root of the tonnage increases, the time required to load and unload the ship (as reflected by log(time)) also increases. The strength of the relationship also appears to have improved compared to M1. This is due to the transformations depicting the data-points being spread more evenly throughout the x-axis. However, past approximately the 85 unit mark of sqrt(tonnage), more data-points would be preferred. Overall, M2 fits the data better than M1, being more reliable in the high-tonnage ranges but could still use a few more data-points. 

    *Answer text here*

(e) (5 marks) Provide the model summary and diagnostics checking plots for model $M_2.$ Does the straight line regression model $M_2$ seem to fit the data well? Comment about the output in a few concise sentences.

    **Answer:**

Model Summary: 

The fitted model equation is: log(Time) = 2.2104 + 0.02151 × sqrt(Tonnage)  

Slope interpretation (β̂₁ = 0.02151): For every unit increase in the square root of tonnage, the log of time is expected to increase by an average of 0.0215 units.

Intercept interpretation (β̂₀ = 2.2104) When the tonnage is zero, the logarithm of time is expected to be 2.2104. This means the time is approximetly 9.12 units. This measurement could be the for the time the ship spends in port doing other activities such as refueling or stocking food for the crew.

The p-values for both coefficients are both less than 0.001. This provides strong evidence that rejects the null hypothesis. Therefore, it is suggested that there is a significant relationship between the square root of tonnage and logarithm of time.


Diagnostic Plots Analysis:

Residuals vs Fitted: The plot depicts no clear patterns. Therefore, the relationship is suggested to be reasonably linear.

Q-Q Residuals: Most of the observations lie along the diagonal, except a few datapoints at the extremities of the graph. This suggests the residuals are generally linear.

Scale-Location: There is no clear pattern visable but that may be due to the limited amounts of data past the 4.0 range of fitted values. Based off the data given, it is suggested that the variance is constant. (Idk if this conclusion is right pls check thx)

Residuals vs Leverage: Again we are looking for data points in the graph with both a high leverage and a standardized residual far from zero. For the most part all data points follow that trend except data point 3, which has a slightly low standardized residual value for it being larger in leverage. Hence further investigation of data point 3 is reccomended.


    *Answer text here*

(f) (4 marks) Perform a hypothesis testing for a positive slope at a significance level of 5% based on model $M_2.$

    **Answer:**

    *Answer text here*

(g) (6 marks) Compare a 95% confidence interval of the mean response and a 95% prediction interval for a new value when Tonnage = 10,000 using the untransformed model $M_1$ and transformed model $M_2$ respectively. Provide two scatterplots that consist the fitted model, the confidence and prediction intervals for each of $M_1$ and $M_2$ respectively. Comment about the output in a few concise sentences.

    **Answer:**

    *Answer text here*
