---
title: "Assignment 1 Submission"
author: "Ian Chen, Marco Gunawan"
date: "10/04/2025"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Statement of Contribution

Include this if pairing.

State what each team member has contributed to the assignment before answering the questions. Each student must contribute towards ALL questions. Your assignment will not be marked without this statement.

```{r eval=FALSE, include=FALSE}
## There are a few options in every chunk, click the setting button (greyed) near the triangle:
# 1. to use default as in the above (show code and output, ie ECHO=TRUE) or 
# 2. to show output only (ie ECHO=FALSE) or
# 3. show nothing (run code) ie include=FALSE
# 4. show nothing (dont run the code), like this chunk, ie eval=FALSE, include=FALSE
## You can knit this template into Word and then update it into pdf etc..
## You can create your own way for submission, this is just an example if you are going to use Rmd.
```

# Question 1 Normal distribution. (8 marks)

(a) (1 mark) Find the probability that the user spends more than 15 minutes per month at the site.

    **Answer:** 0.9938

    ```{r}
    1 - pnorm(15, mean = 25, sd = 4)
    ```

(b) (2 marks) Find the probability that the user spends between 20 and 35 minutes per month at the site.

    **Answer:** 0.8881

    ```{r}
    pnorm(35, mean = 25, sd = 4) - pnorm(20, mean = 25, sd = 4)
    ```

(c) (2 marks) What is the amount of time per month a user spends on Facebook, if only 1% of users spend this time or longer on Facebook?

    **Answer:** 34.3054

    ```{r}
    qnorm(0.99, mean = 25, sd = 4)
    ```

(d) (3 marks) Between what values do the time spent of the middle 90% distribution of Facebook users fall?

    **Answer:** (18.42059, 31.5794)

    ```{r}
    qnorm(0.05, mean = 25, sd = 4)
    qnorm(0.95, mean = 25, sd = 4)
    ```

# Question 2 Blood fat concentration (11 marks)

(a) (6 marks) Conduct a two-independent sample $t$-test using R to determine whether the concentration of plasma cholesterol is significantly different between patients with no evidence of heart disease and those with narrowing of the arteries.

    **Answer:**

    1.  Hypotheses:

        $H_0: \mu_1 = \mu_2$ against $H_A:  \mu_1 \ne  \mu_2$

        where $\mu_1$ and $\mu_2$ are the population means of plasma cholesterol concentrations in patients with no evidence of heart disease and those with narrowing of the arteries respectively.

    2.  Assuming unequal variances, the test statistic $t$ can be given as:

        $$
        t = \frac{195.2745-216.1906}{\sqrt{\frac{1303.9231}{51} + \frac{1850.2488}{320}}} = -3.7357
        $$

        Similarly, the degrees of freedom $df$ is given by:

        $$
        df = \frac{(\frac{1303.9251}{51}+\frac{1850.2488}{320})^2}{\frac{1}{51-1}(\frac{1303.9231}{51})^2+\frac{1}{320-1}(\frac{1850.2488}{320})^2} = 74.5745
        $$

    3.  the sampling distribution thus is $t_{df=74.5745}$.

    4.  The calculated p-value = 0.0003641 \< 0.01.

    5.  Decision. Given the p-value is less than the significance level (0.01), we reject the null hypothesis $H_0$ at the 1% significance level

    6.  Conclusion. Therefore, we conclude that the population means of plasma cholesterol concentrations in patients with no evidence of heart disease and those with narrowing of the arteries are significantly different.

    ```{r}
    #No Disease
    mean1 <- 195.2745
    var1 <- 1303.9231
    n1 <- 51

    #Disease
    mean2 <- 216.1906
    var2 <- 1850.2488
    n2 <- 320
    ```

    ```{r}
    #calculate t and df
    t_stat <- (mean1 - mean2)/sqrt(var1/n1 + var2/n2)
    df <- (var1/n1 + var2/n2)^2 / ((var1/n1)^2/(n1-1) + (var2/n2)^2/(n2-1))

    #calculate P_val
    p_val <- 2 * pt(-abs(t_stat), df)
    ```

(b) (3 marks) Determine a 99% confidence interval for the mean difference in concentration of plasma cholesterol between the two groups of patients.

    **Answer:**

    A 99% confidene interval for the mean difference in concentration of plasma cholesterol between the two groups of patients can be given as:

    (-35.7164, -6.1158)

    With the mean_difference = -20.9161, standard_error = 5.5990, critical t_value = 2.6434 and same degrees of freedom from (a), at df = 74.5745 \~ 75.

    ```{r}
    # Calculate the mean difference
    mean_diff <- mean1 - mean2

    # Calculate the standard error of the difference
    se_diff <- sqrt(var1/n1 + var2/n2)

    # Find critical t-value for 99% confidence interval (two sided)
    t_crit <- qt(0.995, df)

    # Calculate the margin of error
    margin_error <- t_crit * se_diff

    # Calculate the confidence interval
    lower_ci <- mean_diff - margin_error
    upper_ci <- mean_diff + margin_error
    ```

(c) (2 marks) Explain the correspondence between the confidence interval in (b) and a test of the hypotheses you listed in question (a).

    **Answer:**

    Since the confidence interval does not contain zero, this aligns with rejecting the null hypothesis at the 1% significance level in question (a).

    The confidence interval is entirely negative, meaning that the mean plasma cholesterol concentration in the group with no heart disease is lower than the mean in the group with narrowing of arteries. The magnitude of the difference falling between 6.12 and 35.72 units with 99% confidence.

    The fact that zero is not in the interval corresponds directly to the p-value (0.0003641) being less than 0.01, both leading to the same conclusion of rejecting $H_0$.

# Question 3 Regression (31 marks)

(a) (2 marks) Fit a simple linear model $M_1$ to these data. Present the appropriate scatterplot and plot the fitted line onto the scatterplot. Comment about the output in a few concise sentences.

    **Answer:**

    ```{r include=FALSE}
    #read .txt file
    glakes <- read.table("glakes.txt", header = TRUE)
    summary(glakes)

    #create the linear model M_1
    glakes_lm_1 <- lm(Time~Tonnage, data = glakes)
    ```

    ```{r echo=FALSE}
    library(ggplot2)
    ggplot(data = glakes) + 
      geom_point(mapping = aes(x = Tonnage, y = Time)) +
      geom_abline(intercept = coef(glakes_lm_1)[1], slope = coef(glakes_lm_1)[2], color = "red", linewidth = 1) +
      theme(legend.position = "bottom") +
      labs(title = "Great Lakes Time over Tonnage",
           x = "Tonnage", 
           y = "Time")
    ```

(b) (5 marks) Provide the model summary and diagnostics checking plots for model $M_1.$ Does the straight line regression model $M_1$ seem to fit the data well? Comment about the output in a few concise sentences.

    **Answer:**

    The model summary for model $M_1$:

    ```{r echo=FALSE}
    #model summary
    summary(glakes_lm_1)
    ```

    Diagnostics checking plots for model $M_1$

    ```{r echo=FALSE}
    # diagnostics checing plots
    par(mfrow = c(2,2))
    plot(glakes_lm_1)
    ```

(c) (5 marks) Do you think there are outliers or influential points in the data? What influence do these points have on the model fit? Use leverage and Cook's distance for this investigation.

    **Answer:**

    Having a look at the Cook's distance plot:

    ```{r echo=FALSE}
    par(mar = c(4, 4, 1.5, 1.5))
    plot(glakes_lm_1, which = 5, cook.levels = 4 / (nrow(glakes) - 2))
    abline(h = 2, col = "red", lty = 2)   # horizontal line at y = 2
    abline(h = -2, col = "red", lty = 2)  # horizontal line at y = -2
    ```

    We say that a point is classified as a leverage point in simple linear regression, if it's leverage $h_{ii}$ \> 4/$n$. As $n$ = 31, the rule is $h_{ii}$ \> 0.13. As leverage $h_{ii}$ in the plot above has two points with $h_{ii}$ \> 0.13, there are two leverage points in the data.

    We say that points are outliers if their standardized residuals have an absolute value of greater than 2. From the plot above, we can see that there are 2 points whose standardized residuals have an absolute value of greater than 2, such that there are 2 outliers.

    Influential points are observations that have a big influence on the fitted line. This happens when the observations are both high leverage and is unusual (potentially an outlier). We note that the only point which meets both of these criterias is observation 3 and fitting a model after removing this point gives us:

    ```{r echo=FALSE}
    glakes_clean <- glakes[-c(3), ]
    glakes_lm_1_clean <- lm(Time~Tonnage, data = glakes_clean)

    # Create a small data frame with model info
    lines_df <- data.frame(
      intercept = c(coef(glakes_lm_1)[1], coef(glakes_lm_1_clean)[1]),
      slope = c(coef(glakes_lm_1)[2], coef(glakes_lm_1_clean)[2]),
      Model = c("Original", "Cleaned")
    )

    ggplot(data = glakes) + 
      geom_point(mapping = aes(x = Tonnage, y = Time)) +
      geom_abline(data = lines_df, 
              mapping = aes(intercept = intercept, slope = slope, color = Model),
              linewidth = 1) +
      scale_color_manual(values = c("Original" = "red", "Cleaned" = "blue")) +
      theme(legend.position = "bottom") +
      labs(title = "Great Lakes Time over Tonnage",
       x = "Tonnage", 
       y = "Time",
       color = "Model")
    ```

    The new model better fits what appear to be a linear relationship

(d) (4 marks) Fit a regression model to the transformed $M_2$ model. Present the appropriate scatterplot and plot the fitted line onto the scatterplot. Does the transformed line regression model $M_2$ seem to fit the data well? Comment about the output in a few concise sentences.

    **Answer:**

    ```{r}
    #model M_2 data transformations
    log_Time = log(glakes$Time)
    sqrt_Tonnage = sqrt(glakes$Tonnage)

    #linear regression model M_2
    glakes_lm_2 <- lm(log_Time~sqrt_Tonnage, data = glakes)
    ```

    ```{r echo=FALSE}
    #plot transformed M_2 model
    ggplot(data = glakes) + 
      geom_point(mapping = aes(x = sqrt_Tonnage, y = log_Time)) +
      geom_abline(intercept = coef(glakes_lm_2)[1], slope = coef(glakes_lm_2)[2], color = "red", linewidth = 1) +
      theme(legend.position = "bottom") +
      labs(title = "",
           x = "sqrt(Tonnage)", 
           y = "Log(Time)")
    ```

(e) (5 marks) Provide the model summary and diagnostics checking plots for model $M_2.$ Does the straight line regression model $M_2$ seem to fit the data well? Comment about the output in a few concise sentences.

    **Answer:**

    ```{r echo=FALSE}
    #model summary
    summary(glakes_lm_2)
    ```

    ```{r echo=FALSE}
    # diagnostics checing plots
    par(mfrow = c(2,2))
    plot(glakes_lm_2)
    ```

    (4 marks) Perform a hypothesis testing for a positive slope at a significance level of 5% based on model $M_2.$

    **Answer:**

    1.  $H_0: \beta_1 \le 0$ against $H_A:  \beta_1 >  0$

        where $\beta_1$ is the slope of the model $M_2$. Note that the alternative hypothesis is one sided.

    2.  The test statistic (from R output as shown in question (e):

        $$
        t = \frac{\hat{\beta}-0}{se(\hat{\beta})}=\frac{0.021514}{0.001909}=11.27
        $$

    3.  The sampling distribution for the test statistic t is $t_{df=(n-2)}$ that is $t_{df=31-2=29}$.

    4.  The p-value = P($t_{29}$ \> 11.27) = 2.048434e-12

        ```{r include = FALSE}
        pt(11.27, 29, lower.tail=FALSE)
        ```

    5.  Decision. Given the p-value is less than the significance level (0.05), we reject the null hypothesis $H_0$ at the 5% significance level.

    6.  Conclusion. Therefore, we conclude that the slope is statistically significantly positive.

(f) (6 marks) Compare a 95% confidence interval of the mean response and a 95% prediction interval for a new value when Tonnage = 10,000 using the untransformed model $M_1$ and transformed model $M_2$ respectively. Provide two scatterplots that consist the fitted model, the confidence and prediction intervals for each of $M_1$ and $M_2$ respectively. Comment about the output in a few concise sentences.

    **Answer:**

    *Answer text here*
