---
title: "Assignment 1 Submission"
author: "Your name(s)"
date: "submission date"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Statement of Contribution

Include this if pairing.

State what each team member has contributed to the assignment before answering the questions. Each student must contribute towards ALL questions. Your assignment will not be marked without this statement.

```{r eval=FALSE, include=FALSE}
## There are a few options in every chunk, click the setting button (greyed) near the triangle:
# 1. to use default as in the above (show code and output, ie ECHO=TRUE) or 
# 2. to show output only (ie ECHO=FALSE) or
# 3. show nothing (run code) ie include=FALSE
# 4. show nothing (dont run the code), like this chunk, ie eval=FALSE, include=FALSE
## You can knit this template into Word and then update it into pdf etc..
## You can create your own way for submission, this is just an example if you are going to use Rmd.
```


# Question 1 Normal distribution. (8 marks) 


(a) (1 mark) Find the probability that the user spends more than 15 minutes per month at the site. 

    **Answer:**

    *Answer text here*

    ```{r}
    # Answer code here
    ```

(b) (2 marks) Find the probability that the user spends between 20 and 35 minutes per month at the site.

    **Answer:**

    *Answer text here*

(c) (2 marks) What is the amount of time per month a user spends on Facebook, if only 1\% of users spend this time or longer on Facebook? 

    **Answer:**

    *Answer text here*


(d) (3 marks) Between what values do the time spent of the middle 90\% distribution of Facebook users fall? 

    **Answer:**

    *Answer text here*


# Question 2 Blood fat concentration (11 marks)

(a) (6 marks) Conduct a two-independent sample $t$-test using R to determine whether the concentration of plasma cholesterol is significantly different between patients with no evidence of heart disease and those with narrowing of the arteries.

    **Answer:**

    *Answer text here*

(b) (3 marks) Determine a 99\% confidence interval for the mean difference in concentration of plasma cholesterol between the two groups of patients.

    **Answer:**

    *Answer text here*

(c) (2 marks) Explain the correspondence between the confidence interval in (b) and a test of the hypotheses you listed in question (a).

    **Answer:**

    *Answer text here*

# Question 3 Regression (31 marks) 

(a) (2 marks) Fit a simple linear model $M_1$ to these data. Present the appropriate scatterplot and plot the fitted line onto the scatterplot. Comment about the output in a few concise sentences.

    **Answer:**

    *Answer text here*

(b) (5 marks) Provide the model summary and diagnostics checking plots for model $M_1.$ Does the straight line regression model $M_1$ seem to fit the data well? Comment about the output in a few concise sentences. 

    **Answer:**

    *Answer text here*

(c) (5 marks) Do you think there are outliers or influential points in the data? What influence do these points have on the model fit? Use leverage and Cook's distance for this investigation. 

    **Answer:**

    *Answer text here*

(d) (4 marks) Fit a regression model to the transformed $M_2$ model. Present the appropriate scatterplot and plot the fitted line onto the scatterplot. Does the transformed line regression model $M_2$ seem to fit the data well? Comment about the output in a few concise sentences.

    **Answer:**

    *Answer text here*

(e) (5 marks) Provide the model summary and diagnostics checking plots for model $M_2.$ Does the straight line regression model $M_2$ seem to fit the data well? Comment about the output in a few concise sentences.

    **Answer:**

    *Answer text here*

(f) (4 marks) Perform a hypothesis testing for a positive slope at a significance level of 5\% based on model $M_2.$

    **Answer:**

    *Answer text here*

(g) (6 marks) Compare a 95\% confidence interval of the mean response and a 95\% prediction interval for a new value when Tonnage = 10,000 using the untransformed model $M_1$ and transformed model $M_2$ respectively. Provide two scatterplots that consist the fitted model, the confidence and prediction intervals for each of $M_1$ and $M_2$ respectively. Comment about the output in a few concise sentences.

    **Answer:**

    *Answer text here*
